Intentional release of native species undermines ecological stability

### 2022-12-24 rmd/manuscript.Rmd
Update manuscript.Rmd

The intentional release, however, can have wide-ranging consequences than previously thought, as the impacts can propagate through a diversity of ecological interactions [@careyNativeInvadersChallenges2012]. This recognition has sparked a discussion of how the massive introduction of native species alters short-term ecosystem dynamics [@hasegawaAreHatcheryChum2018; @nelsonEcologicalImplicationsChanging2019; @cucheroussetAreDomesticatedFreshwater2020]. Yet, current debates overlook the fact that we rarely assessed the community-wide impact in the long-term. Ecological theory suggests that the stable coexistence of competing species through density-dependent feedback underpins the emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Intentional release may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@laikreCompromisingGeneticDiversity2010; @careyNativeInvadersChallenges2012; @kitadaEconomicEcologicalGenetic2018]. Hence, this form of species management may intervene in the ecological process that allows competing species to coexist, ultimately degrading the long-term community stability. Evidence for this hypothesis is lacking, however.

@@

Two unique strengths in our dataset may have helped uncover the qualitative agreement between theory and empirical patterns. First, human activities are strictly regulated in the protected watersheds (**Figure S10**). Any form of exploitation, including angling, is prohibited in these areas. In addition, exclusive permission from the governer of Hokkaido is required to physically alter in-stream and riparian habitats. Therefore, unmeasured human influences should have minimal influences on the observed relationships. As such, we were able to confine potential confounding factors to natural environmental variation, which were properly accounted in our statistical model (see **Methods**). Second, the sampling method in this monitoring program (a combination of cast net and electrofishing) is carefully chosen to capture stream fishes with different microhabitat preference (e.g., water-column vs. bottom). Thus, our data is robust to potential bias caused by sampling methods. Nevertheless, our results must be viewed with some caution: we cannot exclude the possibility of spurious correlations due to the inherent nature of correlative analysis and field research. Since an experimental approach is nearly impossible at this spatial scale, the application of novel causal inferences (e.g., convergent cross mapping) [@sugiharaDetectingCausalityComplex2012; @ushioFluctuatingInteractionNetwork2018; @suzukiDecomposingPredictabilityIdentify2022] may be a potential alternative to confirm the causal relationship between intentional release and community dynamics. Unfortunately, these methods were not applicable to our data because they require a longer and non-sparse time series to yield robust results.

>>>>

The intentional release, however, can have wide-ranging consequences than previously thought, as the impacts can propagate through a diversity of ecological interactions [@careyNativeInvadersChallenges2012]. This recognition has sparked a discussion of how the massive introduction of native species alters short-term ecosystem dynamics [@hasegawaAreHatcheryChum2018; @nelsonEcologicalImplicationsChanging2019; @cucheroussetAreDomesticatedFreshwater2020]. Yet, current debates overlook the fact that we have rarely assessed the community-wide impact in the long-term. Ecological theory suggests that the stable coexistence of competing species through density-dependent feedback underpins the emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Intentional release may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@laikreCompromisingGeneticDiversity2010; @careyNativeInvadersChallenges2012; @kitadaEconomicEcologicalGenetic2018]. Hence, this form of species management may intervene in the ecological process that allows competing species to coexist, ultimately degrading the long-term community stability. Evidence for this hypothesis is lacking, however.

@@

Two unique strengths in our dataset may have helped uncover the qualitative agreement between theory and empirical patterns. First, human activities are strictly regulated in the protected watersheds (**Figure S10**). Any form of exploitation, including angling, is prohibited in these areas. In addition, exclusive permission from the governer of Hokkaido is required to physically alter in-stream and riparian habitats. Therefore, unmeasured human influences should have minimal influences on the observed relationships. As such, we were able to confine potential confounding factors to natural environmental variation, which were properly accounted in our statistical model (see **Methods**). Second, the sampling method in this monitoring program (a combination of cast net and electrofishing) is carefully chosen to capture stream fishes with different microhabitat preference (e.g., water-column vs. bottom). Thus, our data is robust to potential sampling bias. Nevertheless, our results must be viewed with some caution: we cannot exclude the possibility of spurious correlations due to the inherent nature of correlative analysis and field research. Since an experimental approach is nearly impossible at this spatial scale, the application of novel causal inferences (e.g., convergent cross mapping) [@sugiharaDetectingCausalityComplex2012; @ushioFluctuatingInteractionNetwork2018; @suzukiDecomposingPredictabilityIdentify2022] may be a potential alternative to confirm the causal relationship between intentional release and community dynamics. Unfortunately, these methods were not applicable to our data because they require a longer and non-sparse time series to yield robust results.
